CS 51 Final Project Technical Specification: Matrix Module and Simplex Algorithm---{\annot{Annotated!}}

### 2013-05-05 FinalSpec_Annotated.tex

We have decided to reduce the amount of functions available to the outside
world. The matrix module will still contain the same functionality as before,
but only the functions necessary for the implementation of the Simplex algorithm
will be exposed (row-reduction, for example{\annot{Row Reduction ended up not
being necessary. We have instead exposed a large set of library functions and,
additionally, some functions specific to Simplex that would not be exposed otherwise.
All of this is documented in MatrixI.ml}}). The main focus of the project has
now shifted to having a working simplex algorithm. Other functions, such as
finding the eigenvalues of a matrix, while interesting, have been placed in our
“Cool Features” category. Furthermore, following the idea that Simplex is our
main focus, additional functionality such as parsing input, providing arbitrary
float precision and/or providing a matrix implementation with bignums has been
degraded in importance. All of our attention will be focused on creating a
working simplex algorithm. {\annot{All of this happened, except for the parsing input part. We were able to parse inputs from plaintext files}}.

@@

Our matrix module will actually be a functor that takes in an
\verb@ORDERED_AND_OPERATIONAL@ module. The \verb@ORDERED_AND_OPERATIONAL@ module
will be defined as is below. There will be a global order for the elements, also
as defined below. In our beta implementations, we plan to simply use Ocaml
floats as the elements, but, if time allows, the plan is to make the simplex
algorithm rely on abstract as matrices a possible so that we can change the
\verb@ORDERED_AND_OPERATIONAL@ to something more efficient and/or accurate.

\begin{verbatim}
type order = Equal | Less | Greater

@@

The functions in the module, since they will be dedicated to matrices and
optimized for matrix elements, will not be exposed to the outside world. The
module will be passed in as a constructor to the Matrix module (defined later),
which will exposed only that while is, once again, necessary for the Simplex
module. Additionally, the \verb@ORDERED_AND_OPERATIONAL@ module will more than
likely have additional values and functions not exposed in its signature. For
example, we plan to define the “epsilon for comparison” for floats inside the
module, but we keep it hidden simply because it's not necessary anywhere else.

Users will have no direct access to this module, and therefore there are not
standards that need to be enforced (other than those in place by the interface).

@@

\begin{verbatim}
module type MATRIX =
sig
exception NonSquare
exception ImproperDimension

@@

{\annot{We were able to implement inverse, transpose, trace, and determinant.
The first 3 functions we were able to implement as specified, but the
determinant was implemented by decomposing it into upper and lower triangular
matrices}}

>>>>

We have decided to reduce the amount of functions available to the outside
world. The matrix module will still contain the same functionality as before,
but only the functions necessary for the implementation of the Simplex algorithm
will be exposed (row-reduction, for example{\annot{Row Reduction ended up not
being necessary. We have instead exposed a large set of library functions and,
additionally, some functions specific to Simplex that would not be exposed otherwise.
All of this is documented in MatrixI.ml}}). The main focus of the project has
now shifted to having a working simplex algorithm. Other functions, such as
finding the eigenvalues of a matrix, while interesting, have been placed in our
“Cool Features” category. Furthermore, following the idea that Simplex is our
main focus, additional functionality such as parsing input, providing arbitrary
float precision and/or providing a matrix implementation with bignums has been
degraded in importance. All of our attention will be focused on creating a
working simplex algorithm. {\annot{All of this happened, even the parsing input part. 
We were able to parse inputs from plaintext files}}.

@@

Our matrix module will actually be a functor that takes in an
\verb@ORDERED_AND_OPERATIONAL@ module. The \verb@ORDERED_AND_OPERATIONAL@ module
will be defined as is below. There will be a global order for the elements, also
as defined below. In our beta implementations, we plan to simply use Ocaml
floats as the elements, but, if time allows, the plan is to make the simplex
algorithm rely on abstract as matrices a possible so that we can change the
\verb@ORDERED_AND_OPERATIONAL@ to something more efficient and/or accurate.
{\annot{In the end, we ended up using the Nums OCaml library in order to obtain
arbitrary precison and size with our elements. We made this decision late in the
coding, but because we coded everything abstractly, changing from Floats to Nums
was as simple as exposing the right files and setting the right modules}}

{\annot{Updated version of this can be found in EltsI.ml}}
\begin{verbatim}
type order = Equal | Less | Greater

@@

The functions in the module, since they will be dedicated to matrices and
optimized for matrix elements, will not be exposed to the outside world. The
module will be passed in as a constructor to the Matrix module (defined later),
which will exposed only that while is, once again, necessary for the Simplex
module. Additionally, the \verb@ORDERED_AND_OPERATIONAL@ module will more than
likely have additional values and functions not exposed in its signature. For
example, we plan to define the “epsilon for comparison” for floats inside the
module, but we keep it hidden simply because it's not necessary anywhere else.
{\annot{This was mostly implemented as described.}}

@@

{\annot{This Module changed significantly. For an updated look, see MatrixI.ml}}
\begin{verbatim}
module type MATRIX =
sig
exception NonSquare
exception ImproperDimension

@@

{\annot{We were able to implement inverse, transpose, trace, and determinant.
The first 3 functions we were able to implement as specified, but the
determinant was implemented by decomposing it into upper and lower triangular
matrices. The functions that weren't implemented were mostly left behind because
they were not essential to the Simplex Algorithm}}
